---
title: "STAT 5014 HW7"
author: "Max McGill"
date: '`r Sys.Date()`'
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=T, tidy.opts=list(width.cutoff=55),
                tidy=T, message=F, warning=F)
    library.warn <- library
    library <- function(package, help, pos = 2, lib.loc = NULL, character.only = FALSE,
                        logical.return = FALSE, warn.conflicts = FALSE, quietly = TRUE,
                        verbose = getOption("verbose")) {
       if (!character.only) {
          package <- as.character(substitute(package))
       }
       suppressPackageStartupMessages(library.warn(
          package, help, pos, lib.loc, character.only = TRUE,
          logical.return, warn.conflicts, quietly, verbose))}
    
    fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            text
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

## **Problem 2: Sums of Squares**

### **Part A: For Loop**

```{r, echo=F}
#generate data
set.seed(12345)
y <- rnorm(n = 1e7, mean = 1, sd = 1)

#part a: for loop
SSTa<-0
t1<-system.time({
    ybar<-mean(y)
    for(i in 1:1e7){
        SSTa <- SSTa + (y[i]-ybar)^2
    }
})
knitr::kable(cbind(SSTa,t1["user.self"],t1["sys.self"],t1["elapsed"]),col.names=c('SST','User','System','Elapsed'),caption="Timed For Loop")
```

### **Part B: Vector Operations**

```{r, echo=F}
#Part b: vector operations
SSTb <- 0
t2 <- system.time({
    y_star <- (y-mean(y))
    SSTb <- crossprod(y_star)
    })
knitr::kable(cbind(SSTb,t2["user.self"],t2["sys.self"],t2["elapsed"]),col.names=c('SST','User','System','Elapsed'),caption="Vector Operations")
```

### **Part C: dopar**

It takes too long for my computer. The code is in the Appendix.

```{r, echo=F, eval=F}
#packages
library(foreach)
library(doParallel)
library(parallel)

#Part c: dopar
c1<-makeCluster(2)
registerDoParallel(c1)
system.time(
  SSTc<-foreach(i=1:length(y), .combine='+') %dopar% ((y[i]-ybar)^2)
)
stopCluster(c1)
knitr::kable(cbind(SSTc,t2["user.self"],t2["sys.self"],t2["elapsed"]),col.names=c('SST','User','System','Elapsed'),caption="dopar")
```

### **Part D: parSapply**

It takes too long for my computer. The code is in the Appendix.

```{r, echo=F, eval=F}

#Part d: parSapply
sstfunc<-function(y){
  SSTdv <- c()
  for(i in 1:length(y)){
    SSTd[i] <- (y[i]-mean(y))^2
  }
  return(SSTd)
}
cl <- makeCluster(2)
clusterExport(cl,"sstfunc")
system.time({
    SSTd <- parSapply(cl, 1:length(y),
            function(y) sstfunc,USE.NAMES = F)
})
stopCluster(cl)
knitr::kable(cbind(SSTd,t2["user.self"],t2["sys.self"],t2["elapsed"]),col.names=c('SST','User','System','Elapsed'),caption="parSapply")


```


## **Problem 3: Gradient Descent**

```{r, echo=F}
#object designation
    set.seed(1256)
    theta <- as.matrix(c(1,2),nrow=2)
    X <- cbind(1,rep(1:10,10))
    h <- X%*%theta+rnorm(100,0,0.2)
#more objects
    thetanew<-as.matrix(c(0,0),nrow=2)
    alpha<-0.05
    m<-length(h)
    tol<-0.1

#the computation
for(i in 1:m){
  while(abs(thetanew[1]-theta[1])>tol && abs(thetanew[2]-theta[2])>tol){
    thetanew[1]<-theta[1]-alpha/m*sum(thetanew[1]+thetanew[2]*X[i,2]-h[i])
    thetanew[2]<-theta[2]-alpha/m*sum((theta[1]+theta[2]*X[i,2]-h[i])*X[i,2])  
    }
}
thetanew

#for comparison
lm(h~0+X)
```

Here we have the result obtained from the method, which is followed by the result from using the linear model function.

## **Problem 4: Bootstrapped Regression**

I couldn't get this bootstrap to work in time, having run into problems coordinating the dimensionality and applying parallelization. The code is in the Appendix. 

```{r, echo=F, eval=F}
#from assignment
    set.seed(1267)
    n <- 200
    X <- 1/cbind(1,rt(n,df=1),rt(n,df=1),rt(n,df=1))
    beta <- c(1,2,3,0)
    Y <- X %*% beta + rnorm(100,sd=3)
#bootstrap
B<-10000
Ymat<-matrix(rep(Y,4),4*n,4)
bootbeta<-matrix(0,4*B,4)
for(b in 1:B){
  bootid<-sample(1:n,n,replace=T)
  bootx<-X[bootid,]
  booty<-Ymat[bootid,]
  bootbeta[b]<-coef(lm(booty~bootx))
}
knitr::kable(summary(lm(Y~X))$coefficients, caption="Summary of Bootstrap Model Parameters")
```

## **Appendix 1: R Code**
```{r, eval=F, include=T}
##problem 2
set.seed(12345)
y <- rnorm(n = 1e7, mean = 1, sd = 1)

#part a: for loop
SSTa<-0
t1<-system.time({
    ybar<-mean(y)
    for(i in 1:1e7){
        SSTa <- SSTa + (y[i]-ybar)^2
    }
})
knitr::kable(cbind(SSTa,t1["user.self"],t1["sys.self"],t1["elapsed"]),col.names=c('SST','User','System','Elapsed'),caption="Timed For Loop")

#part b: vector operations
SSTb <- 0
t2 <- system.time({
    y_star <- (y-mean(y))
    SSTb <- crossprod(y_star)
    })
knitr::kable(cbind(SSTb,t2["user.self"],t2["sys.self"],t2["elapsed"]),col.names=c('SST','User','System','Elapsed'),caption="Vector Operations")

#packages
library(foreach)
library(doParallel)
library(parallel)

#Part c: dopar
c1<-makeCluster(2)
registerDoParallel(c1)
system.time(
  SSTc<-foreach(i=1:length(y), .combine='+') %dopar% ((y[i]-ybar)^2)
)
stopCluster(c1)
knitr::kable(cbind(SSTc,t2["user.self"],t2["sys.self"],t2["elapsed"]),col.names=c('SST','User','System','Elapsed'),caption="dopar")

#Part d: parSapply
sstfunc<-function(y){
  SSTdv <- c()
  for(i in 1:length(y)){
    SSTd[i] <- (y[i]-mean(y))^2
  }
  return(SSTd)
}
cl <- makeCluster(2)
clusterExport(cl,"sstfunc")
system.time({
    SSTd <- parSapply(cl, 1:length(y),
            function(y) sstfunc,USE.NAMES = F)
})
stopCluster(cl)
knitr::kable(cbind(SSTd,t2["user.self"],t2["sys.self"],t2["elapsed"]),col.names=c('SST','User','System','Elapsed'),caption="parSapply")

##problem 3
#object designation
    set.seed(1256)
    theta <- as.matrix(c(1,2),nrow=2)
    X <- cbind(1,rep(1:10,10))
    h <- X%*%theta+rnorm(100,0,0.2)
#more objects
    thetanew<-as.matrix(c(0,0),nrow=2)
    alpha<-0.05
    m<-length(h)
    tol<-0.1

#the computation
for(i in 1:m){
  while(abs(thetanew[1]-theta[1])>tol && abs(thetanew[2]-theta[2])>tol){
    thetanew[1]<-theta[1]-alpha/m*sum(thetanew[1]+thetanew[2]*X[i,2]-h[i])
    thetanew[2]<-theta[2]-alpha/m*sum((theta[1]+theta[2]*X[i,2]-h[i])*X[i,2])  
    }
}
thetanew

#for comparison
lm(h~0+X)

##problem 4
#from assignment
    set.seed(1267)
    n <- 200
    X <- 1/cbind(1,rt(n,df=1),rt(n,df=1),rt(n,df=1))
    beta <- c(1,2,3,0)
    Y <- X %*% beta + rnorm(100,sd=3)
#bootstrap
B<-10000
Ymat<-matrix(rep(Y,4),4*n,4)
bootbeta<-matrix(0,4*B,4)
for(b in 1:B){
  bootid<-sample(1:n,n,replace=T)
  bootx<-X[bootid,]
  booty<-Ymat[bootid,]
  bootbeta[b]<-coef(lm(booty~bootx))
}
knitr::kable(summary(lm(Y~X))$coefficients, caption="Summary of Bootstrap Model Parameters")
```

